High-Performance Language Primitives for Distributed Architectures using Asynchronous Events

### 2012-07-19 doc/asplos2013/intro.tex
Working on the intro

\section{Introduction}
\label{sec:intro}
Introduction goes here\cite{Keeneland}.

>>>>

\section{Introduction}
\label{sec:intro}

% Big machines -> latencies are growing
% Processors can't be blocked, need to be able to continue handling work
% Operations need to be composable
% deferred execution model

% Contributions:
% - An event system where every operation can take an event and returns an event
% - Handles are globally valid
% - More flexible synchronization primitive for this environment: deferred locks
% - Physical regions for data movement including special versions for reductions

With the multicore revolution, modern supercomputers have continued to scale
by increasing the number of processors available for computation.  The
price of this scaling is complex interconnection networks and
deep memory hierarchies.  These parts of the system are designed to
minimize latency, but the growing scale of these machines has
caused average and worst-case latencies to grow by orders of magnitdue.
Recently, some of these latencies have grown to the point that supercomputers
share many of the same performance properties as machines commonly
classified as distributed systems.  Despite this major evolution in the
performance of supercomputers, most programs for supercomputers are still
written using the same old programming models and libraries that were designed
years ago for machines many orders of magnitude smaller.

In this paper we propose a new interface for programming modern supercomputers 
that is designed to hide the large latencies associated with current 
interconnection networks and memory hierarchies.  Our interface is built 
around {\em events}.  Events provide a mechanism for naming a point in the future
when an operation (i.e. task, copy, synchronization) will complete.  Every call to perform
an operation in our runtime is asynchronous and returns immediately with an event that will 
{\em trigger} when the operation has completed.  Furthermore, every call
to perform an operation can take a precondition event that must trigger before
the operation can begin.  Using events programmers can compose large
chains of dependent operations.  The runtime is free to execute operations
in any order that obeys the event dependences.  With this freedom the
runtime can schedule operations in a way that optimizes throughput
while automatically hiding the long latency memory and communication operations.

For most supercomputer programmers effectively overlapping computation with
memory and communication latency is the most difficult barrier to achieving high
performance.  Current parallel programming APIs hinder this ability by either
failing to support asynchronous constructs, or by supporting constructs that are
not composable.  For example, MPI supports asynchronous send and receive operations
but does so in a way that later requires blocking to know when the send and receive 
operations have completed\cite{MPI}.  There is no way in MPI to launch a new 
operation contingent on the result of a blocking receive operation.

% Parallel programs have three parts
%  Parallel control
%  Data movement in the memory hierarchy
%  Synchronization
